---
title: "Initial Korean covid 19"
author: "Jonathan Urbano"
date: "June 21, 2023"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## METHODS

We are gonna analyze the sex of the patients and the result if they are alive or dead.I will use the variables in my collection of cat~cat. This is a type of problem that we can use the helper file and we will copy into our project to get the correct answer.





## Graphical results - use ggplot commands to do this  


#### Plot barplots ggplot way

```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

In the counting chart i can see that a lot people died. It is 50-70. In the percentage chart, female is 35 and male is about 63. There are more females than males, however twice as many males are dying in percentage.

  





  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

In the percentages i can see that more females are alive with the difference of 12%. Of all people that are dead 36% is female and 64% are male. In the second chart i see that female are 98.7 alive and male are only 97. In the dead % female were 1.4% and male 3%.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In the group of 1831female were alive and 1411male were alive. All in all more female are alive than male. The p value is 1 out of a thousand in the Chi-square. It is a small chance to get male and female dying. In fisher's test 9 out of 10,000 the are about the same p value. You can also see the odds ratio of. the odds of female lving against the odds of male dying. We want to know that the odds of this is happening.


## Conclusions
In today's lesson we learned to do a ggplot with the helper file. We changed the class for result and name the project kRdata. We were also able to create a numerical graph in % and counting as well as run a Chi-square test and a fisher test which both are equal in p value. The final result is that we want to know what the odds are of this happening. Males are the ones who suffered more during this study, hence females had an advantage over male as a result.

